---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# shinygouv

<!-- badges: start -->
[![R-CMD-check](https://github.com/spyrales/shinygouv/workflows/R-CMD-check/badge.svg)](https://github.com/spyrales/shinygouv/actions)
<!-- badges: end -->

The goal of {shinygouv} is to ...

## Installation

You can install the development version of shinygouv from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("spyrales/shinygouv")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(shinygouv)
## basic example code
```


## Code of Conduct

Please note that the shinygouv project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
