---
title: "Creating a project"
author: "Sebastien Rochette"
date: "23/01/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r development, include=FALSE}
library(testthat)
```

<!--
**Developers**

This allows to create a directory ready to be proposed to user of your package, while including 

- the necessary parts of your "renv.lock"
- a README for your users to follow instructions
- your vignettes

This flat file contains a function that you will be using for your deliveries.  
*This is not recommended to use it for a RStudio New project template Menu, because this would imply for your users to install your package twice, with bad dependencies.*

Delivery?

At the end of this file, you will see the code that you can use for each delivery.

What to do?

=> Follow instructions written as HTML comments like the following one

- Create a vignette in your package that will be used as a Readme for your users project
  + I recommend using the template in {thinkridentity}
-->

```{r dev-first-time}
# Empty vignette
# usethis::use_vignette('e-user-project-readme')

# {thinkridentity} template
file.copy(
  system.file("package", "e-user-project-readme.Rmd", package = "thinkridentity"),
  here::here("vignettes"), 
  recursive = TRUE
)
```

<!--
- Search and replace the "the-vignette.Rmd" by a vector of your vignettes to include in the project for your users to test. Here and in the vignette for readme.
- Search and replace the name of 'the.package' by the name of your package. Here and in the vignette for readme
-->

```{r development-update}
# To modify ----
# Which vignettes to copy for users to test the package
the_vignettes <- c("the-vignette.Rmd")
# Name of the vignette used as Readme
the_readme_vignette <- "e-user-project-readme.Rmd"

# If rsproject does not exist ----
if (!dir.exists(here::here("inst/rsproject"))) {
  dir.create(here::here("inst/rsproject"), recursive = TRUE)
}

# If specific versions need to be forced ----
# renv::install("usethis@2.0.0")
# renv::install("odbc@1.2.2")

# Load already included functions
pkgload::load_all(export_all = FALSE)

# Update DESCRIPTION ----
attachment::att_amend_desc(
  # pkg_ignore = c("whereami"), # if needed
  # extra.suggests = c("DT", "knitr") # if needed
)

# Update renv in the current development project ----
custom_packages <- c(
  attachment::att_from_description(),
  "renv",
  "devtools", "roxygen2", "usethis", "pkgload",
  "testthat", "covr", "attachment", "remotes",
  "bookdown", "DT", "pak",
  "zip", "credentials",
  "gert",
  "fontawesome",
  "pkgdown", "styler", "checkhelper", "fusen", "boxr"
)
renv::snapshot(packages = custom_packages, prompt = FALSE)

# Create dir with RStudio project ready ----
dir.create(here::here("inst/rsproject"))
dir.create(here::here("inst/rstudio/templates/project"), recursive = TRUE)

# Copy Rprofile
file.copy(here::here(".Rprofile"), here::here("inst/rsproject/Rprofile"), overwrite = TRUE)

# Copy renv files
dir.create(here::here("inst/rsproject/renv"))
file.copy(here::here("renv/.gitignore"), here::here("inst/rsproject/renv/.gitignore"), overwrite = TRUE)
file.copy(here::here("renv/activate.R"), here::here("inst/rsproject/renv/activate.R"), overwrite = TRUE)
file.copy(here::here("renv/settings.dcf"), here::here("inst/rsproject/renv/settings.dcf"), overwrite = TRUE)

# Upgrade renv.lock in the user project ----
renv::install("usethis@1.6.3")

custom_packages <- c(
  attachment::att_from_description(field = c("Depends", "Imports", "Suggests")),
  "renv", "pkgload",
  "remotes", "styler", "knitr", "rmarkdown", "rstudioapi",
  "DT", "zip", "credentials",
  "gert",
  "fontawesome", "pak"
)
renv::snapshot(packages = custom_packages, lockfile = here::here("inst/rsproject/renv.lock"), prompt = FALSE)

lock_ <- renv:::lockfile(file = here::here("inst/rsproject/renv.lock"))

# Enable this universe
options(
  repos = c(
    # thinkropen = 'https://thinkr-open.r-universe.dev',
    CRAN = "https://cran.rstudio.com"
  )
)
lock_$repos(
  # thinkropen = 'https://thinkr-open.r-universe.dev',
  CRAN = "https://cran.rstudio.com"
)

lock_$write(file = here::here("inst/rsproject/renv.lock"))
rm(lock_)

# Update vignettes ----
for (vignette in the_vignettes) {
  file.copy(here::here("vignettes", vignette), 
            here::here("inst/rsproject", vignette), overwrite = TRUE)
}

# Update Readme and clean yaml ----
temp_readme <- tempfile("readme", fileext = ".Rmd")
file.copy(here::here("vignettes", the_readme_vignette),
          temp_readme)
readme_lines <- readLines(temp_readme)
w.clean <- grep("WARNING", readme_lines)[1]
w.yaml.end <- grep("^---$", readme_lines)[2]
readme_lines <- readme_lines[-c(1:max(w.clean, w.yaml.end))]
cat(c("---
output: github_document
editor_options: 
  chunk_output_type: console
---", 
readme_lines), 
sep = "\n", 
file = here::here("inst/rsproject/README.Rmd"),
append = FALSE)


# Back to normal Rprofile ----
source(here::here(".Rprofile"))
# Re-install development versions if needed
# renv::install("usethis@2.0.0")
```


<!--
From here, if you inflate this flat file,
this can be used to create a RStudio template for new project.
However, this is a bad idea because this would require your users to install your package twice: on their general installation, and in the {renv} environment of the project created.

-- So you can stop here --

Or change "the-vignette.Rmd" by a vector of your vignettes included
-->

## Create directory with everything needed to use the package

You can create a new directory with Rmd files needed to use the package, as well as the {renv} environment to reproduce the conditions of development.  

### Create the project with RStudio or command line

- Use the RStudio project creation with the template 'the.package'
    - File > New Project > New directory > "Admin project with {the.package}"
    - Choose where to create the project
    
OR with command line

- Run `the.package::create_new_project("your/path/of/choice")`
  - Restart the R session inside the new project

### Restore the conditions of development

- To install all R packages with correct versions inside the newly created package, run:

```r
renv::restore()
```

- Re-install the {the.package} package inside this environment

```r
# install.packages("remotes")
remotes::install_local("the.package_0.1.0.tar.gz", upgrade = "never")
```

### Use the package

Then, you can use the files:  

- "the-vignette.Rmd" to test the package

```{r function-create_new_project}
#' Create directory with everything needed to use the package
#'
#' @param path Path of the directory where to create the project
#' @param open Logical. Whether to open the directory when created
#' @param overwrite Logical. Whether to overwrite all existing files.
#'
#' @importFrom utils browseURL
#' @importFrom usethis create_project
#' @export
create_new_project <- function(path = tempfile("path"), open = FALSE,
                             overwrite = FALSE) {
  path <- normalizePath(path, mustWork = FALSE)
  if (!dir.exists(path)) {
    dir.create(path)
  } else {
    message("Files added to existing path: ", path)
  }
  
  files_to_copy <- file.path(
    app_sys_admin("rsproject"), 
    list.files(app_sys_admin("rsproject")))
  files_to_copy <- files_to_copy[!grepl("Rprofile", files_to_copy)]
  
  file.copy(from = files_to_copy, to = path,
            recursive = TRUE, overwrite = overwrite)
  
  # Specific to .Rprofile
  file.copy(from = app_sys_admin("rsproject", "Rprofile"),
            to = file.path(path, ".Rprofile"),
            overwrite = overwrite)
  
  # Create the project 
  create_project(path, open = open)
  
  if (isTRUE(open)) {
    browseURL(path)
  }
  
  return(path)
}
```

```{r examples-create_new_project}
# Define a directory, new or existing
my_directory <- tempdir()
# Create the content
create_new_project(my_directory)
# See the content of the directory
fs::dir_tree(my_directory)
```

```{r tests-create_new_project}
test_that("create_new_project works properly", {
  
  skip_if_not(Sys.getenv("RSTUDIO_PROGRAM_MODE") == "desktop")
  
  # Define a directory, new or existing
  my_directory <- tempfile(pattern = "admin-")

  # Create the content
  create_new_project(my_directory, open = FALSE)

  # fs::dir_tree(my_directory, all = TRUE)

  all_files <- c("renv.lock", ".Rprofile", "renv/activate.R",
                 "README.Rmd",
                 "the-vignette.Rmd")

  for (the_file in all_files) {
    expect_true(file.exists(file.path(my_directory, the_file)))
  }

  # No error to open file, but message because of existing directory
  expect_message(create_new_project(my_directory, open = TRUE), regexp = "Files added to existing path")


# Test locally during dev only that content was updated ----
  message("create_new_project is up-to-date")
  skip_if_not(file.exists("../../.Rprofile"))
  
  all_files <- c(".Rprofile", "renv/activate.R") # "renv.lock", 
  for (the_file in all_files) {
    file_expected <- readLines(file.path("..", "..", the_file))
    file_created <- readLines(file.path(my_directory, the_file))
    expect_equal(file_created, file_expected)
  }
  
  all_files <- c("the-vignette.Rmd")
  for (the_file in all_files) {
    file_expected <- readLines(file.path("..", "..", "vignettes", the_file))
    file_created <- readLines(file.path(my_directory, the_file))
    expect_equal(file_created, file_expected)
  }
  
})
```


```{r development-inflate, eval=FALSE}
# Run but keep eval=FALSE to avoid infinite loop
# Execute in the console directly
fusen::inflate(flat_file = "dev/flat_userproject.Rmd", 
               vignette_name = "d - create admin project", 
               open_vignette = FALSE,
               overwrite = "yes",
               check = FALSE)

# Test file for updated skeleton
pkgload::load_all()
testthat::test_file(here::here("tests/testthat/test-create_new_project.R"))
```

<!-- 

-- Delivery --

- Run the content of the `development-update` chunk above to update the renv.lock
- Run the code in the `development-delivery` chunk below to
  + Create a directory that will be send to your users
  + Build sources of your package to include in the delivery
  + Include the sources


-->

```{r development-delivery}
# _RS Project update ----
rs_dir <- "deliverables/the.package_project"
unlink(rs_dir, recursive = TRUE)
usethis::use_git_ignore("deliverables")
pkgload::load_all()
outdir <- create_new_project(path = file.path(tempdir(), "the.package_project"))
file.copy(outdir, "deliverables", recursive = TRUE)
unlink(outdir, recursive = TRUE)
# => Do not forget to add sources in the project

# _Sources build ----
temp_install <- file.path(tempdir(), "temp_pkg_folder")
unlink(temp_install, recursive = TRUE)
dir.create(temp_install)

# Build with all except what is in Rbuildignore
rbuildignore <- gsub("\\\\[.]", "[.]", paste(readLines(".Rbuildignore"), collapse = "|"))
all.files.noinst <- list.files(here::here())[
  !grepl(rbuildignore, list.files(here::here()))]

# Copy the entire folder in tempdir()
file.copy(from = all.files.noinst,
          to = temp_install,
          recursive = TRUE) 
list.files(temp_install)

# Build sources from reduced package
dir.create("deliverables/sources")
devtools::build(pkg = temp_install,
                path = "deliverables/sources",
                vignettes = FALSE) # build the package

unlink(temp_install, recursive = TRUE)

# _Copy sources in the.package_project ----
# dir.create("deliverables/the.package_project/renv/cellar", recursive = TRUE)
file.copy(tail(
  list.files("deliverables/sources", pattern = ".tar.gz", full.names = TRUE), 1),
          to = "deliverables/the.package_project", overwrite = TRUE)

# Verify the Readme, but you maybe deliver this "deliverables/the.package_project" as is to your clients
```

